---
title: "Demographic Influences on Climate Change Perceptions in the Midwest"
subtitle: "EAES 480 Final Presentation"
author: "Dylan Janowski"
institute: "University of Illinois at Chicago"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(infer)
library(knitr)
library(xaringanthemer)
library(broom)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.asp = 0.6
)
```

```{r load-data, include=FALSE}
data <- read.csv("../data/data.csv")
```

```{r}
xaringanthemer::style_xaringan(
  text_color = "#222222",
  header_color = "#1b5e20",
  background_color = "#ffffff",
  link_color = "#2e7d32",
  title_slide_background_color = "#2e7d32",
  title_slide_text_color = "#ffffff"
)
```

```{r}
names(data) <- c(
  "climate_happening",
  "human_cause",
  "planet_impact",
  "seriousness",
  "conserve_resources",
  "env_vs_econ",
  "human_rights_resources",
  "regulations_too_far",
  "gov_action",
  "corporate_responsibility",
  "individual_impact",
  "support_regulations",
  "support_carbon_tax",
  "who_pays_cost",
  "coal_tradeoff",
  "personal_behavior",
  "pay_more_sustainable",
  "largest_climate_driver",
  "tech_solution",
  "age",
  "education",
  "community_current",
  "community_grew_up"
)

data <- data %>%
  mutate(across(everything(), as.factor))
```

# How Do Demographics Shape Climate Change Beliefs?

**Why This Matters**
- Because public opinion shapes environmental action, understanding how climate beliefs vary across demographic groups is important for improving communication and addressing climate change effectively. To study this, I conducted a survey within the Midwest asking various climate focused questions to respondents.

**Guiding Questions**
- Does belief in human causation differ by current community?
- Does belief in human causation differ by age group?
- Are these differences statistically significant?

**Demographics of Focus**
- Current community
- Age group
---

# Data Overview

- Survey dataset of **262 respondents**
- **23 variables** (demographics + opinions)
  - **Key variables:**
      - **human_cause** (Do you believe climate change is primarily caused by human activity?)
      - **community_current** (Which best describes the community where you currently live?)
      - **age** (Age range)

---

# Data Wrangling
.pull-up[
```{r, eval=FALSE, echo=T}
names(data) <- c(
  "climate_happening",
  "human_cause",
  "planet_impact",
  "seriousness",
  "conserve_resources"
# ... Renaming continued

data <- data %>%
  mutate(across(everything(), as.factor))
```
]

.pull-down[
```{r, echo=T}
data %>%
  select(human_cause, community_current, age) %>%
  slice(c(1, 6, 10))
```
]

---

# Does belief in human causation differ by current community?

.pull-left[
```{r, eval=FALSE, echo=T}
ggplot(data, aes(x = community_current, fill = human_cause)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c(
    "Yes" = "#4CAF50",
    "No" = "#E57373",
    "Maybe" = "#FFD54F"
  )) +
  labs(
    title = "Belief that Climate Change is Human-Caused by Community",
    x = "Community Type",
    y = "Proportion",
    fill = "Climate Belief"
  ) +
  theme_minimal()
```
]

.pull-right[
```{r}
ggplot(data, aes(x = community_current, fill = human_cause)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c(
    "Yes" = "#4CAF50",
    "No" = "#E57373",
    "Maybe" = "#FFD54F"
  )) +
  labs(
    title = "Belief that Climate Change is Human-Caused by Community",
    x = "Community Type",
    y = "Proportion",
    fill = "Climate Belief"
  ) +
  theme_minimal()
```
]

---

# Are These Community Differences Statistically Significant?

.pull-up[
**Null hypothesis:**
Community and belief in human-caused climate change have no relationship

**Alternative hypothesis:**
Community and belief in human-caused climate change have a relationship

**Test Results**
- Because p < 0.05, we reject the null hypothesis
- There is strong statistical evidence that climate change beliefs depend on community type
]

.pull-down[
```{r}
chisq.test(table(data$community_current, data$human_cause))
```
]

---

# Does belief in human causation differ by age group?

.pull-left[
```{r, eval=FALSE, echo=T}
ggplot(data, aes(x = age, fill = human_cause)) +
  geom_bar(position = "fill") +
  scale_x_discrete(limits = c("Under 20", "20-30", "30-40", "40-50", "50-60", "60+")) +
  scale_fill_manual(values = c(
    "Yes" = "#4CAF50",
    "No" = "#E57373",
    "Maybe" = "#FFD54F"
  )) +
  labs(
    title = "Belief that Climate Change is Human-Caused by Age",
    x = "Age Group",
    y = "Proportion",
    fill = "Human Cause Belief"
  ) +
  theme_minimal()
```

]

.pull-right[
```{r}
ggplot(data, aes(x = age, fill = human_cause)) +
  geom_bar(position = "fill") +
  scale_x_discrete(limits = c("Under 20", "20-30", "30-40", "40-50", "50-60", "60+")) +
  scale_fill_manual(values = c(
    "Yes" = "#4CAF50",
    "No" = "#E57373",
    "Maybe" = "#FFD54F"
  )) +
  labs(
    title = "Belief that Climate Change is Human-Caused by Age",
    x = "Age Group",
    y = "Proportion",
    fill = "Human Cause Belief"
  ) +
  theme_minimal()
```
]

---

# Does belief in human causation differ by age group?

.pull-left[
```{r, echo=T}
data <- data %>%
  mutate(
    age_group = case_when(
      age == "Under 20" | age == "20-30" |
        age == "30-40" ~ "40 and under",
      age == "40-50" | age == "50-60" |
        age == "60+" ~ "40 and above",))
```
]

.pull-right[
```{r}
ggplot(data, aes(x = age_group, fill = human_cause)) +
  geom_bar(position = "fill") +
  scale_x_discrete(limits = c("40 and under", "40 and above")) +
  scale_fill_manual(values = c(
    "Yes" = "#4CAF50",
    "No" = "#E57373",
    "Maybe" = "#FFD54F"
  )) +
  labs(
    title = "Belief that Climate Change is Human-Caused by Age Group",
    x = "Age Group",
    y = "Proportion",
    fill = "Human Cause Belief"
  ) +
  theme_minimal()
```
]

---

# Are these age differences statistically significant?

.pull-left[

```{r, echo=T}
set.seed(480)
n_boot <- 1000
boot <- replicate(n_boot, {
  sample_data <- data %>%
    slice_sample(n = nrow(data), replace = TRUE)

  sample_data %>%
    filter(age_group == "40 and above") %>%
    summarise(p_yes = mean(human_cause == "Yes", na.rm = TRUE)) %>%
    pull(p_yes)
})

mean(boot)
quantile(boot, c(0.025, 0.975))
```
]

--

.pull-right[

```{r}
ggplot(data.frame(boot), aes(x = boot)) +
  geom_histogram(bins = 30, fill = "#4CAF50", color = "white") +
  labs(
    title = "Bootstrap Estimate of Climate Belief Ages 40+",
    x = "Proportion Saying 'Yes'",
    y = "Bootstrap Frequency"
  ) +
  theme_minimal(base_size = 14)
```

- The models shows 49% of respondents over 40 believe climate change is human-caused.
- However, the quantile range is 38% to 59%.
]

---

# Are these age differences statistically significant?

.pull-left[
```{r, echo=T}
set.seed(480)
n_boot <- 1000
boot_under40 <- replicate(n_boot, {
  sample_data <- data %>%
    slice_sample(n = nrow(data), replace = TRUE)

  sample_data %>%
    filter(age_group == "40 and under") %>%
    summarise(p_yes = mean(human_cause == "Yes", na.rm = TRUE)) %>%
    pull(p_yes)
})

mean(boot_under40)
quantile(boot_under40, c(0.025, 0.975))
```

]

--

.pull-right[
```{r}
ggplot(data.frame(boot_under40), aes(x = boot_under40)) +
  geom_histogram(bins = 30, fill = "#4CAF50", color = "white") +
  labs(
    title = "Bootstrap Estimate of Climate Belief Ages 40 and Under",
    x = "Proportion Saying 'Yes'",
    y = "Bootstrap Frequency"
  ) +
  theme_minimal(base_size = 14)
```
- The models shows 84% of respondents over 40 believe climate change is human-caused.
- However, the quantile range is 78% to 89%.
]

---

# Conclusions

- There is statistically significant evidence that community context is associated with differences in belief in human-caused climate change.
- Age also shows variation, with younger and older groups differing in both level of agreement and uncertainty.
- Bootstrap results suggest that while most groups lean toward believing climate change is human-caused, confidence ranges are wide in some community and age groups, indicating uncertainty in estimates.
- Overall, demographic factors help explain how environmental beliefs are distributed, which has implications for how climate communication and policy support are shaped.
