---
title: "Project proposal"
author: "Dylan Janowski"
output: html_document
---

> GM GRADE: 18/20 - great work! Some additional narrative to guide the reader through the sequence of figures and your interpretation is missing, and some more concrete details about how the variables breakdown into response and predictor, how they are related to the questions, and # responses by group would have strenthened.

> Description: 7.5/8
> Content: 9/10
> Quality: 1.5/2

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Data

This data contains information of demographic factors such as age, education, and community and how they relate to perceptions of environmental questions. The data came from a survey of 262 individuals collected in public and university settings, with each row representing one respondent and each column (23 of them) capturing answers to survey questions and demographic information.

```{r, message = FALSE}
data <- read.csv("../data/data.csv")

names(data) <- c(
  "climate_happening",
  "human_cause",
  "planet_impact",
  "seriousness",
  "conserve_resources",
  "env_vs_econ",
  "human_rights_resources",
  "regulations_too_far",
  "gov_action",
  "corporate_responsibility",
  "individual_impact",
  "support_regulations",
  "support_carbon_tax",
  "who_pays_cost",
  "coal_tradeoff",
  "personal_behavior",
  "pay_more_sustainable",
  "largest_climate_driver",
  "tech_solution",
  "age",
  "education",
  "community_current",
  "community_grew_up"
)

```

I renamed the variables (questions) to shorter, easier names to manipulate and visualize later.

```{r}
head(data)
```

## 2. Exploratory Visualization

```{r}
ggplot(data, aes(x = community_current, fill = climate_happening)) +
  geom_bar(position = "fill") +
  labs(
    title = "Belief that Climate Change is Happening by Community",
    x = "Community Type",
    y = "Proportion",
    fill = "Belief"
  )
```

```{r}
ggplot(data, aes(x = community_current, fill = seriousness)) +
  geom_bar(position = "fill") +
  labs(
    title = "Seriousness of Climate Change by Community",
    x = "Community Type",
    y = "Proportion",
    fill = "Seriousness"
  )
```

```{r}
ggplot(data, aes(x = community_current, fill = human_cause)) +
  geom_bar(position = "fill") +
  labs(
    title = "Belief that Climate Change is Human Caused by Community",
    x = "Community Type",
    y = "Proportion",
    fill = "Human Cause Belief"
  )
```

```{r}
ggplot(data, aes(x = community_grew_up, fill = climate_happening)) +
  geom_bar(position = "fill") +
  labs(
    title = "Belief that Climate Change is Happening by Community Grew up in",
    x = "Community Grew Up In",
    y = "Proportion",
    fill = "Belief"
  )
```

```{r}
ggplot(data, aes(x = community_grew_up, fill = seriousness)) +
  geom_bar(position = "fill") +
  labs(
    title = "Seriousness of Climate Change by Community Grew up in",
    x = "Community Grew Up In",
    y = "Proportion",
    fill = "Seriousness"
  ) + 
  scale_fill_viridis_d() + 
  theme_classic()
```

```{r}
ggplot(data, aes(x = community_grew_up, fill = human_cause)) +
  geom_bar(position = "fill") +
  labs(
    title = "Belief that Climate Change is Human Caused by Community Grew up in",
    x = "Community Grew Up In",
    y = "Proportion",
    fill = "Human Cause Belief"
  )
```

```{r}
ggplot(data, aes(x = age, fill = human_cause)) +
  geom_bar(position = "fill") +
  scale_x_discrete(limits = c("Under 20", "20-30", "30-40", "40-50", "50-60", "60+")) +
  labs(
    title = "Belief that Climate Change is Human-Caused by Age",
    x = "Age Group",
    y = "Proportion",
    fill = "Human Cause Belief"
  ) 
  ggsave("figure1.png", height = 5 , width = 8)
```
Tried to reorder the graph in increasing age by converting to factor data, but was having trouble with errors, so I instead found that I can manually reorder it.

## 3. Possible question and method

As of now, I'm unsure of what I should focus on with the data. Should I focus on one short question such as "do urban/rural/suburban individuals disagree on climate change occurrence and its cause," or should I explore wider questions? I will be able to use methods such as t-test and ANOVA tests to compare mean climate attitudes across different demographic groups. I will be able to use these visuals and tests to compare the differemces between the groups and how they compare on answers.
